---
title: "Assignment 1 - Reproducible Research"
author: "Ignacio Gonzálvez"
date: "Saturday, October 15, 2015"
output: html_document
---



## Loading and preprocessing the data

```{r}
temporal <- tempfile()
url <- "https://d396qusza40orc.cloudfront.net/repdata/data/activity.zip"
download.file(url, temporal, method="libcurl")
datos <- read.csv(unz(temporal, "activity.csv"), header=TRUE, sep=",")
unlink(temporal)
# summary(datos)
```
1. Calculate the total number of steps taken per day
```{r}
steps.date <- aggregate(steps ~ date, data = datos, sum)
```
2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day

```{r}
hist(steps.date$steps, col="green", main="Number of steps taken per day", xlab="steps", ylab="frecuency")
```

3. Calculate and report the mean and median total number of steps taken per day
```{r}
Mediadepasosdiaria <- mean(steps.date$steps)
Mediadepasosdiaria
Medianadepasosdiaria <- median(steps.date$steps)
Medianadepasosdiaria
```

## What is the average daily activity pattern?

   1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
    
```{r}
steps.intervalo <- aggregate(steps ~ interval, data = datos, mean)
plot(steps.intervalo, type = "l", xlab="interval [in 5min]", ylab="Average daily activity pattern of steps",  main="Average number of steps")
```

   2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
      
```{r}
Maximoenintervalo <-steps.intervalo$interval[which.max(steps.intervalo$steps)]
Maximoenintervalo
```
Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

    1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
    
    
```{r}
steps.NA <- is.na(datos$steps)
sum(steps.NA)
```

    2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
    I put the mean instead of NA
    
```{r}
datos <- merge(datos, steps.intervalo, by = "interval", suffixes = c(".x", ".y"))
datos$steps[NAS] <- datos$steps.y[NAS]
datos <- datos[, c(1:3)]

```

# Perform the imputation
```{r}
for (i in 1:nrow(datos)){
  if (is.na(datos$steps[i])){
    intervalo_valores <- datos$interval[i]
    row_id <- which(steps.intervalo$interval == intervalo_valores)
    steps_val <- steps.intervalo$steps[row_id]
    datos$steps[i] <- steps_val
  }
}


```

    3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
table_interval_steps <- aggregate(steps ~ interval, datos, mean)
```
    4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?



```{r}
#vector_datos <- c(datos$steps)
#print("Generating the plot1.png file .....")
#png(file="./plot1.png", width = 480, height = 480, units = "px") # open file png

#hist(vector_datos, col="red", main="Number of steps taken per day", xlab="Number of steps")
#dev.off() ## close de png file 
```
